---
title: "Motor Trend Regression Report"
author: "Alex Petkovski"
date: "Sunday, June 21, 2015"
output: 
    pdf_document:
        number_sections: yes
        toc: false
---

#Executive Summary
The following report analyses whether automatic or manual transmission improve MPG
based on data extracted from 1974 Motor Trend US magazine and quantifies this 
difference. We found that manual performed better than automatic, when holding horse 
power and car weight fixed.   

#Data Exploration
```{r, PackagesUsed, echo=FALSE, message=FALSE, comment=FALSE}
library(datasets)
library(ggplot2)
library(GGally)
library(car)
```
After loading required packages (see Appendix), we construct a pairs plot in order 
to identify correlations by parameters shown in **mtcars**. Treating **Transmission(am)** 
as a factor variable shows boxplots and histograms for each possible parameter, 
which from initial inspection reveals that manual transmission generally has a 
higher mpg than automatic transmission when not factoring any other variables. The 
panel plot is shown in the Appendix. Next, we need to set factor variables for all 
factor variables (See Appendix).

```{r, ggPairs, echo=FALSE, comment=FALSE, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.keep='none'}
mtcars$am <- as.factor(mtcars$am)
g = ggpairs(mtcars,
            colour = "am",
            lower = list(continuous = "smooth"),
            upper = list(params=list(size = 2, fatten=0.3)),
            params = c(method = "loess"), 
            axisLabels = "internal",
            title = "Exploratory Pairs Panel Plot by Transmission")
g
```
```{r, ProcessFactors, echo=FALSE, message=FALSE}
mtcars$am <- as.factor(mtcars$am)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
```
#Model Selection
As a consequence of the Panel Plot, we probably won't need qsec and gear due to 
low correlation with mpg. Also, from the model, we see highly correlated parameters 
likely potential to cause Variance Inflation.   

##All parameter model
To show this we fit a model with all parameters which showed high p-values and high
variance inflation factors, thus our model requires a selection process. (See 
appendix for code and results)
```{r, VarianceInflation, echo=FALSE, message=FALSE, results='hide', cache=TRUE, warning=FALSE, fig.keep='none'}
par(mfrow = c(2, 2))
fit.all <- lm(mpg ~ ., mtcars)
summary(fit.all)$coef
```
Results in appendix show non-significant **p-values** for each beta.

#Diagnostics
With each iteration we run diagnostics to determine if model has good R^2^ and low
**VIF**.   

##Initial Model using Transmission Type Only
```{r, InitialFit, echo=FALSE, message=FALSE}
par(mfrow = c(2, 2))
fit <- lm(mpg ~ am, mtcars)
fit1 <- fit
summary(fit1)$coef
intercept <- summary(fit1)$coef[1]
am1 <- summary(fit1)$coef[2]
est_am1 <- intercept + am1
r.squared <- summary(fit1)$r.squared
```
From this model:   
1. Estimate for mpg is `r round(intercept,2)` when transmission is automatic(am 
= 0) and `r round(est_am1,2)` when transmission is manual(am = 1)   
2. We also see that t values are sigmificant, both at less than 0.001   
3. Without taking any other parameters other that transmission type, we see manual 
has better mpg than automatic, but we R-squared is at `r round(r.squared,2)`, so 
our model is biased and requires additional explanatory variables to make any reasoable 
conclusion   

## Add Horse Power (hp) to model
```{r, SecondFit, echo=FALSE, message=FALSE, cache=TRUE}
par(mfrow = c(2, 2))
fit2 <- update(fit, mpg ~ am + hp)
coef2 <- summary(fit2)$coef
r.squared.2 <- summary(fit2)$r.squared
```
1. Coefficients have good p-values 
2. This model with Horse Power added improves R^2^ to `r round(r.squared.2,2)` 
3. VIFs are low 
```{r, SecondFitVif, echo=FALSE, message=FALSE, cache=TRUE}
vif(fit2)
```
## Add weight (wt) to model along with interaction on transmission
```{r, ThirdFit, echo=FALSE, message=FALSE, cache=TRUE}
fit3 <- update(fit, mpg ~ am + hp + wt + wt:am)
summary(fit3)$coef
r.squared.3 <- summary(fit3)$r.squared
```
1. Coefficients continue to have good p-values
2. This model with Weight and interaction added improves R^2^ to `r round(r.squared.3,2)` 
3. VIFs are worse but still low
```{r, ThirdFitVif, echo=FALSE, message=FALSE, cache=TRUE}
vif(fit3)
```
4. Let's plot the residuals as this is our best model (See appendix for plot)
```{r, ThirdFitPlot, echo=FALSE, message=FALSE, cache=TRUE, fig.keep='none'}
par(mfrow = c(2, 2))
plot(fit3)
```
5. Residual plots don't show any obvious bias (See appendix for plot)   

#Conclusion
```{r, Conclusion, echo=FALSE, message=FALSE, cache=TRUE}
anova(fit1, fit2, fit3)
```
From ANOVA, we see that inclusion of all am, hp, and am:wt gives a model with high
significance and low p-value. Best model from the three tested is fit3 and is significant
in the alpha = 0.01 range. Finally, keeping hp and wt constant, manual transmission 
has higher mpg by `r round(sum(summary(fit3)$coef[1:5])-summary(fit3)$coef[1]-sum(summary(fit3)$coef[3:4]),2)`. The table of standard errors are shown in 3.1 coefficients table.

\newpage 

#APPENDIX 
## R Code used in the Study: 
The following code was used to create and represent the data used in this analysis
```{r, PackagesUsed, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following code was used to create the panel plot described in Data Exploration 
section and the panel plot is shown
```{r, ggPairs, echo=TRUE, message = FALSE, warning= FALSE, cache=TRUE}
```
The following code was used to set factor variables
```{r, ProcessFactors, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code and coefs for the all parameter model
```{r, VarianceInflation, echo=TRUE, message=FALSE}
```
The following shows code used to create the initial fit with mpg ~ am
```{r, InitialFit, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code used to create the second fit adding hp
```{r, SecondFit, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code used to create the second fit VIFs
```{r, SecondFitVif, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code used to create the third fit adding wt and am:wt
```{r, ThirdFit, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code used to create the third fit VIFs
```{r, ThirdFitVif, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```
The following shows code used to create the third fit residual plots
```{r, ThirdFitPlot, echo=TRUE, message = FALSE, comment = FALSE, cache=TRUE}
```
The following shows code used to create the ANOVA conclusion results
```{r, Conclusion, echo=TRUE, message = FALSE, comment = FALSE, eval=FALSE}
```